Interfering with Noninterference \iftext\thanks{\bcp{Identify applicable funding agencies here (DARPA and NSF and...?)}}\fi

### 2020-04-09 main.tex
Writeup of eager stuff safety

A {\em trace} is a finite or infinite sequence of machine states,
$\seq{\mach} = \mach_0,\, \mach_1,\ldots$, in which adjacent machine
states are related by $\stepstounderfull$---i.e., in which
$\mach_{i-1} \stepstounderfull \mach_i$ for each $i > 0$ in the
sequence.

@@

A pair of traces $\seq{M}$ and $\seq{N}$ satisfy the {\em
  confidentiality} safety hyperproperty with respect to a contour
$\contour$, if, if the initials states are variants $N_0 \in
\variant{C}{M_0}$, then their observations coincide:
$\obstrace(\contour, \seq{M}) = \obstrace(\contour, \seq{N})$.
\leo{Should we cut at minimum length for termination-insensitiveness?}
\bcp{Or just ask that one be a prefix of the other?}

>>>>

A {\em trace} is a finite or infinite sequence of machine states,
$\seq{\mach} = \mach_0,\, \mach_1,\ldots$, in which adjacent machine
states are related by $\stepstounderfull$---i.e., in which
$\mach_{i-1} \stepstounderfull \mach_i$ for each $i > 0$ in the
sequence. We write $\seq{\mach}^i_j$ for the subtrace between $i$ and
$j$.

A {\em contour} $\contour \in \CONTOURS = \COMPONENTS \rightarrow
\{\HIGHSEC, \LOWSEC\} \times \{\HIGHINT, \LOWINT\}$ is a map from
machine components to a pair of confidentiality and integrity labels.
$\LOWSEC$ marks locations whose contents represent data observable
{\em by the current process}; $\HIGHSEC$ corresponds to non-observable
data (e.g. local state of the process caller). $\HIGHINT$ marks
locations that the current process does not have write access to;
$\LOWINT$ is freely accessible memory. As examples of different
combinations consider the following:
\begin{itemize}
\item Locations that belong to the current process such as
  available registers, local stack, memory accessible through
  available pointers will be tagged $\LOWSEC \times \LOWINT$:
  they can be both read and written.
\item The local state of the caller that needs to be protected from a
  potentially untrusted callee will be marked $\HIGHSEC \times
  \HIGHINT$. The callee should be able to neither read nor write
  these locations.
\item The instruction memory is read only for every process: $\LOWSEC
  \times \HIGHINT$.
\item We won't have examples of secret data that can be overwritten.
  \leo{Maybe in the lazy setting?}
\end{itemize}

\subsection{Identity colorings}
\leo{Feel free to coin another name}

An {\em identity coloring} of a trace $\seq{\mach}$ is an equal-length
sequence $\seq{\ID}$ of booleans $\IDS = \{\ME, \NOTME\}$, where
$\ID_j = \ME$ iff the machine state $\mach_j$ is executing the same
process as $\mach_0$.

A {\em call map} is a set of instructions for a machine state that are
(1) $JAL$s and (2) correspond to actual calls.

Given a machine trace $\seq{\mach}$ and a call map this can be
computed as follows:
\begin{enumerate}
\item $mach_0$ is colored $\ME$.
\item Find the next machine state $\mach_c$ (if any) that appears in
  the trace $\seq{mach}$ such that the instruction executed at
  $\mach_c$ is a $JAL$ that is marked as a call in the call map.
\item Color all states between $\mach_0$ and $\mach_c$ inclusive as
  $\ME$. If no such state exists, color the remainder of the trace
  $\ME$.
\item Find the next state $\mach_r$ after $\mach_c$ in the trace (if
  any) such that the pc of $\mach_r$ is equal to the pc of $\mach_c +
  4$, and whose stack pointer is identical to that of $\mach_c$.
\item Color all states between $\mach_{c+1}$ and $\mach_r$ as
  $\NOTME$.  If no such state exists, color the remainder of the trace
  $\NOTME$.
\item Color $\mach_{r+1}$ as $\ME$. Go to step (2).
\end{enumerate}

\subsection{Contoured Trace}

Given a contour $C$ and a machine state $M$, the {\em contoured trace}
induced by $C$ and $M$, written $\induced{C}{M}$, is the (finite or
infinite) sequence of pairs of (contour, machine state) pairs in which
states in adjacent pairs are related by $\stepstounderfull$, and in which
contours in adjacent pairs are constructed as follows:
\begin{itemize}
\item If in the identity coloring of $\seq{\mach}$ $\ID_i$ = $\ID_{i+1}$ then
  $C_{i+1} = C_i$.
\item If $\ID_{i} = \ME$ and $\ID_{i+1} = \NOTME$, then $C_{i+1}$ is
  constructed from $C_i$ by setting the local state of $\mach_i$ (the
  caller) to $\HIGHSEC \times \HIGHINT$.
\item If $\ID_{i} = \NOTME$ and $\ID_{i+1} = \ME$, then $C_{i+1}$ is
  constructed from $C_i$ by setting the local state of the caller
  to $\LOWSEC \times LOWINT$.
\end{itemize}
\leo{To make this more precise we need to accurately describe what
  local state {\em is}}

\subsection{Integrity}

A trace $\seq{\mach}$ satisfies the {\em integrity} safety property
with respect to a contour $\contour$, if for all components
$\component$ where $\contour(\component) = (\_, \HIGHINT)$,
$\mach_i(\component) = \mach_0(\component)$ for all $i$ in the trace.

\leo{For confidentiality we need observations first}

A {\em component observation} is either a word or the special value
$\bot$.  \leo{Fix notation, clashes with $\LOWINT$}  A {\em state
  observation} is a map from machine components to component
observations.

In particular, the observation $\obstrace(\contour, \seq{\mach})$
arising from a pair of a contour $\contour$ and a trace $\seq{\mach}$
is sequence of state observations $\stateobs_0,\, \stateobs_1,\,
\ldots$ defined informally as follows:
\begin{itemize}
\item $\stateobs_0$ maps every visible component $\component$ (where
  $\contour(\component) \in \{\LOWSEC \times \_\}$) to its value in
  $\mach_0$ and every hidden component to $\bot$;
\item $\stateobs_{i+1}$ is defined from $\stateobs_{i}$ by (1) for
  each visible component $\component$ (where
  $\contour(\component) \in \{\LOWSEC \times \_ \}$) its value
  from $\mach_{i+1}$, and (2) for each hidden component $\component$,
  where $\contour (\component) = \{\HIGHSEC \times \_\}$,
  copying its value from $\stateobs_{i}$, and (3) if the resulting
  $\stateobs_{i+1}$ is identical to $\stateobs_{i}$, then skipping
  $\stateobs_{i+1}$.  \iftext\bcp{Too informal!!}\fi
\end{itemize}
\iftext\bcp{Emphasize that ``observability is not observable.''}\fi

\subsection{Confidentiality}

Given a machine state $\mach$ and a contour $\contour$, a {\em
  variant} of $\mach$ with respect to $\contour$ is any machine state
$\mach'$ whose components agree with $\mach$ on all components marked
$(\LOWSEC, \_)$ by $\contour$---i.e., if $\contour(\component) =
(\LOWSEC, \_)$, then $\mach(\component) = \mach'(\component)$.

A pair of traces $\seq{M}$ and $\seq{N}$ satisfy the {\em
  confidentiality} safety hyperproperty with respect to a contour
$\contour$, if, if the initials states are variants of each other $N_0
\in \variant{C}{M_0}$, then the observations of one is a prefix of the
observations of the other: $\obstrace(\contour, \seq{M}) \subseteq
\obstrace(\contour, \seq{N})$.

\subsection{Stack Safety}

A trace $\seq{\mach}$ satisfies {\em stack safety} with respect to a
contour $\contour$ if:
\begin{itemize}
\item $\seq{\mach}$ satisfies integrity with respect to $\contour$.
\item For every variant $N \in \variant{C}{\mach_0}$, $\seq{\mach}$
  and the the induced trace $\seq{N}$ satisfy confidentiality with
  respect to $\contour$.
\item Every subtrace $\seq{\mach}^r_c$ colored $\NOTME$ in the
  identity coloring of $\seq{\mach}$ satisfies stack safety with
  respect to the contour $\contour_c$.
\end{itemize}

\leo{OLD STUFF BEGINS HERE}

@@

A pair of traces $\seq{M}$ and $\seq{N}$ satisfy the {\em
  confidentiality} safety hyperproperty with respect to a contour
$\contour$, if, if the initials states are variants $N_0 \in
\variant{C}{M_0}$, then their observations coincide:
$\obstrace(\contour, \seq{M}) = \obstrace(\contour, \seq{N})$.
\leo{Should we cut at minimum length for termination-insensitiveness?}
\bcp{Or just ask that one be a prefix of the other?}
\bcp{Or go back to halting states.  (Which I guess really means go back to
  Halting + Failing states.  We never quite worked out how that would go.)}
